\subsection{Introduction}

\subsection{First release}
March 11th 2014 marked the first release of recur. It was posted to the fitness subreddit of the popular image board site Reddit\citep{reddit:2005}. The idea was to do a small release to pick up on any bugs that got through testing. It became very helpful because as the developer of the site knew how everything worked because they had built it, other people said they did not know where to click. This subject had been raised by colleague and fixed with an image of a chain link to denote a link. After this feedback had been submitted, changes were made to the title links. They were made more obvious when rolled over with a background colour change and also, if it is the users first visit to the site then they will be shown a message telling them that the titles of the cards are links to more information. Other feedback included users receiving error flashes when really the database was successfully updated. This was because the error flash was getting called if the log entry was saved or not. This caused uses to submit multiple entries to the database for the same activity. This then caused another problem of not being able to removed a logged activity after it was put into the database. An obvious feature was overlooked during development but was soon implemented in to make sure users could delete unwanted logged activities. The way the log was displayed was also rewritten to make it easier to see which data went with which activity. Users also wanted to be able to see which goals they had completed on the dashboard. This was limited down to goals that had been completed in the last 2 weeks. 


\subsection{Conclusion}