\section{Introduction}
This chapter will cover the testing and evaluation of application. It will go into detail about the choices made and give reasons as to why certain tools or methodologies were used over others. It will also discuss the first release and reviewed release of the application, and talk about the user's feedback on each release.\\

\section{Methodologies}
For the main part of testing, test driven development(TDD) was used throughout the building of the application. TDD was used because of the constant deployments to Heroku. All of the features implemented had to be working perfectly before deployment. However, sometimes this was not the case. All of the tests are written so sometimes you miss a test case and this can lead to deploying broken code. Usually, most of the bugs were picked up either with the testing or quite quickly after deployment if any bugs got through. TDD fitted in well with the Agile development approach that was taken when building the application. TDD and Agile development were used to make testing easy. TDD tests the code as you are working so you are not left doing all the test last, or after every feature. This allowed to get working prototypes out and deployed very quickly.\\

\section{Tools}
\subsection{Rspec}
Rspec is a RubyGem that is used to write test cases for test driven development. As test driven development was the chosen methodology, it was important to get the right testing tool for the job. Rspec was chosen for its easy to read and write test cases. Rspec's syntax is designed to human readable. With commands like \lstinline{should have_content} or \lstinline{should_not have_link} it makes writing and debugging the test much easier. Rspec can be coupled with other gems like FactoryGirl to make testing even easier. FactoryGirl \citep{factorygirl:2013} was used to create fake users so testing could go further with by interacting with a web page using another gem called Capybara \citep{capybara:2013}. Capybara allowed interaction with the webpage so we could navigate between pages and see if the correct flashes were showing. All three gems were chosen on their power of being able to test almost anything, their simple syntax which is almost like readable English, and they way they all work seamlessly together.

\subsection{Chrome Development Tools}
Chrome Developer Tools \citep{cdt:2014} was a key part to the testing of the layout of this web application. Using the developer tools you are able to target specific parts of the application and see where they are getting their CSS styles from. This was especially useful when reusing class selectors on the cards. The cards would share similar styles but also have some individual styles that would conflict if kept under one class. It was also very good for mobile testing. It also allows the user to edit the CSS selectors and their values, without reloading the page or changing any files. This proved useful for finding ways to fix certain layout issues as it does not save the changes. The Chrome Developer Tools has an emulation mode, which allows the user to see what the website would look on a selection of devices. This proved more useful than just resizing the browser window as it took viewport into account. \\

\section{Testing}
Test were written before the majority of code. This is under the test driven development methodology. You write a test that you know will fail, and then implement the fix. A test case was written for any foreseeable possibility a user could enter. As original functionality of the application changed, more test cases were implemented.\\

While running the test, it actually brought up something that was not intended. It brought up that some of the methods of retrieving data from the database had been depreciated. Depreciation is a common downside to using frameworks that have multiple versions. Keeping on top of depreciation is very important when it comes to production application because calling a depreciated method can bring your application to a halt.

\section{Evaluation}
\subsection{First release}
March 11th 2014 marked the first release of Recur. It was posted to the fitness subreddit of the popular image board site Reddit\citep{reddit:2005}. The idea was to do a small release to pick up on any bugs that got through testing. It proved very useful because the developer will understand how the site works because they build it but users said they did not know where to click when first visiting the site. This subject had been raised by another user and fixed with an image of a chain link to denote a link. After this feedback had been submitted, changes were made to the title links. They were made more obvious when rolled over with a background colour change and also, if it is the user's first visit to the site then they will be shown a message telling them that the titles of the cards are links to more information. Other feedback included users receiving error flashes when really the database was successfully updated. This was because the error flash was getting called if the log entry was saved or not. This caused users to submit multiple entries to the database for the same activity. This then caused another problem of not being able to removed a logged activity after it was put into the database. An obvious feature was overlooked during development but was soon implemented in order to make sure users could delete unwanted logged activities. The way the log was displayed was also rewritten to make it easier to see which data went with which activity. Users also wanted to be able to see which goals they had completed on the dashboard. This was limited down to goals that had been completed in the last 2 weeks.

\section{Design changes}
Throughout the progression of development, although nothing major has changed, the design has changed in a number of ways for a number of different reason. Probably the biggest aspect that has changed during development was the colour of the website. Originally, it was a shade of blue with some red in it to create a dark lavender colour. It has now been changed to a shade of blue with more green in as opposed to red. The change was purely aesthetic but I feel it was the better choice. The small chart on the front page was dropped quite early on during development. It was very had to implement such a stripped down chart that did not display very much information upon the quick glances it would receive. It was swapped out with a much large, more descriptive chart on the dedicated weights page. Now the chart is one more click away but it allows the chart more space to display more information and therefore became more useful to the user.

\section{Conclusion}
Using Rspec and its accompanying Gems was a great help when it came to testing. Tests were very simple to read and write, and when they picked up any bugs it was easy to identify which test it failed on and what should have happened. Also, using test driven development allowed the application to be deployed very early on because each feature was tested as it was being developed, when it was finished it was ready to be deployed. It also saved a lot of time at the end of development, and saved time in the overall development because each test case was thought of before writing any code, and because each test case was thought of beforehand, it could be taking into account when coding the feature.